A Formal Model of Checked C

### 2023-01-03 sandp23/appendix.tex

\section{Appendix}

@@

\subsection{The Compilation Rules}\label{appx:comp1}
Fig.~\ref{fig:syntaxerased} and Fig.~\ref{fig:semanticserased} shows
the syntax for $\elang$, the target language for compilation. We syntactically
restrict the expressions to be in A-normal form to simplify the
presentation of the compilation rules. In the Redex model, we
occasionally break this constraint to speed up the performance of
random testing by removing unnecessary let bindings.
To allow explicit runtime
checks, we include $\ebounds$ and $\enull$ as part of $\elang$
expressions which, once evaluated, result in an corresponding error state. $\eassignstack{x}{\dot a}$ is a new syntactic form that
modifies the stack variable $x$ with the result of $\dot a$. It is
essential for bounds widening. $\leq$ and $-$ are introduced
to operate on bounds and decide whether we need to halt with a bounds error
or widen a null-terminated string.

\begin{figure}[h]
{\small
  \[\begin{array}{llcl}
      \text{Atoms}       & \dot a & ::= & n \mid x \\

      \text{C-Expressions} & \dot c & ::= & \dot a \mid \estrlen{\dot a}  \mid \emalloc {\dot  a}  \mid \ecall{f} {\overline{\dot a}}  \\
                         & & \mid &  \mid  \dot a \circ \dot a \mid \estar{\dot a} \\
                         & & \mid & \eassign{\dot a}{\dot a} \mid \eassignstack{x}{\dot a}  \mid \eif{\dot a}{\dot e}{\dot e} \\
                         & & \mid & \ebounds \mid \enull \\
      \text{Expressions}  & \dot e & ::= & \dot c \mid \elet{x}{\dot c}{\dot e}\\
      \text{Binops} & \circ & ::= & + \mid - \mid \leq  \\
      \text{Closure} & \dot C & ::= & \hole \mid \elet{x}{\dot a}{\dot C} \\
       & & \mid & \eif{\dot a}{\dot e}{\dot C} \mid \eif{\dot a}{\dot    C}{\dot e} \\
      \text{Bounds Map} & \rho & \in & \texttt{Var} \rightharpoonup \texttt{Var} \times \texttt{Var}
  \end{array}
  \]
}
  \caption{\elang{} Syntax}
  \label{fig:syntaxerased}
{\small
  \[\begin{array}{lll}
    \dot \mu & ::= & n \mid \bot \\
    \dot c & ::= & \ldots  \mid \eret{x}{\dot \mu}{\dot e} \\
    \dot H & \in & \mathbb{Z} \rightharpoonup \mathbb{Z} \\
    \dot r & ::= & \dot e \mid \enull \mid \ebounds \\
    \dot E & ::= & \hole \mid  \elet{x}{\dot E}{\dot e}  \mid \eret{x}{i}{\dot E} \\
             & \mid & \eif{\dot E}{\dot e}{\dot e} \mid \estrlen{\dot E} \\
             & \mid & \emalloc{\dot E} \mid \ecall{f}{\overline{\dot
                      E}} \mid \dot E \circ \dot a \mid n \circ \dot E \\
             & \mid & \estar{\dot E} \mid \eassign{\dot E}{\dot a}
                      \mid \eassign{n}{\dot E} \mid
                      \eassignstack{x}{\dot E} \\
      \overline{\dot E } & ::= & \dot E \mid n,\overline{\dot E} \mid
                                 \overline{\dot E}, \dot a

  \end{array} \]
}
  \caption{\elang Semantic Defs}
  \label{fig:semanticserased}
\end{figure}

$\elang$ does not include any annotations. We remove structs from
$\elang$ because we can always statically convert expressions of the form
$\eamper{n:\tau}{f}$ into $n + n_f$, where $n_f$ is the statically
determined offset of $f$ within the struct. We ellide the semantics of
$\elang$ because it is self-evident and mirrors the semantics
$\lang$. The difference is that in $\elang$, only $\ebounds$ and
$\enull$ can step into an error state. All failed dereferences and
assignments would result in a stuck state and therefore we rely on the
compiler to explicitly insert checks for checked pointers.

Fig.~\ref{fig:compilation} and Fig.~\ref{fig:compilation2}
 shows the rules for the compilation judgment for expressions,
\[\Gamma;\rho \vdash e \gg \dot C, \dot a\]
The judgment is presented differently from the one in
Sec.~\ref{sec:compilation}, which was simplified for presentation purposes. First,
we remove $\Theta$ and $m$ because these
parameters are only used for checking and have no
impact on compilation. Second, the judgment includes two
outputs, a closure $\dot C$ and an atom expression $\dot a$, instead of
a single $\elang$ expression $\dot e$. $\dot C$ can be intuitively understood as a
partially constructed program or context. Whereas $\dot E$ is used for
evaluation, $\dot C$
is used purely as a device for compilation. As an example,
when compiling $(1 : \tint) \plus (2 : \tint)$, 
we would first create a fresh variable $x$, and then produce two outputs:
\[    \dot C = \elet{x}{1 \plus 2}{\hole}\]
\[    \dot a = x\]
To obtain the compiled expression $\dot e$, we plug $\dot a$ into
$\dot C$ using the usual notation $\dot C[\dot a]$. We can also use $\dot C$ to represent runtime checks, which
usually take the form $\elet{x}{\dot c}{\hole}$, where $\dot c$
contains the check whose evaluation must not trigger $\ebounds$ or $\enull$ for the program to
continue (see Fig.~\ref{fig:compilationhelpers} for the metafunctions that create
those checks).

This unconventional output format enables us to separate the
evaluation of the term and the computation that relies on the term's
evaluated result. Since effects and reduction (except for variables) happen only
within closures, we can precisely control the order in which effects
and evaluation happen by composing the contexts in a specific order.
Given two closures $\dot C_1$
and $\dot C_2$, we write $\dot C_1[\dot C_2]$ to denote the meta
operation of plugging $\dot C_2$ into $\dot C_1$. We also use
$\dot C_{a;b;c}$ as a shorthand for $\dot C_a[\dot C_b[\dot C_c]]$. In
the \textsc{C-Ind} rule,
we first evaluate the expressions that correspond to $e_1$ and $e_2$
through $\dot C_1$ and $\dot C_2$, and then perform a null check and
an addition through $\dot C_n$ and $\dot C_3$. Finally, we dereference
the result through $\dot C_4$ before returning the pair $\dot C_4,\dot
x_4$, propagating the flexibility to the compilation rule that recursively
calls \textsc{C-Ind}.

Fig.~\ref{fig:compilationhelpers} shows the metafunctions that create
closures representing dynamic checks. These functions first examine
whether the pointer is a checked. If the pointer is unchecked, an empty closure $\hole$
will be returned, because there is no need to perform a check. For bounds
checking, there is a special case for NT-array pointers, where the
bounds are retrived from the \textcolor{purple}{shadow} variables (found by
looking up $\rho$) on the stack  rather than
using the bounds specified in the type annotation. This is how we
achieve the same precise runtime behavior as $\lang$ in our compiled expressions.

Fig.~\ref{fig:compilationhelpers2} shows the metafunctions related to
bounds widening. $\vdash_{extend}$ takes $\rho$,
a checked NT-array pointer variable $x$, and its bounds $(b_l,b_h)$ as inputs, and returns an
extended $\rho'$ that maps $x$ to two fresh variables $x_l$, $x_h$,
together with a closure $\dot C$ that initializes $x_l$ and $x_h$ to
$b_l$ and $b_h$ respectively. This function is used in the
\textsc{C-Let} rule to extend $\rho$ before compiling the body of the
$\elettext$ binding. The updated $\rho'$ can be used for generating
precise bounds checks, and for inserting expressions that can potentially
widen the upper bounds, as seen in the $\vdash_{widenstr}$
metafunction used in the \textsc{C-Str} compilation rule.

\begin{figure*}[t]
{\small
  \begin{mathpar}
    \inferrule[]
    { x_l, x_h = \fresh \\
      \rho' = \rho[x \mapsto (x_l, x_h)]\\
    \dot C = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}}
  {\cextend{\rho}{x}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot C, \rho'}\\}

  \inferrule[]
  {x_l,x_h = \rho(x) \\
    x_w = \fresh \\
  \dot C = \elet{x_w}{\eif{x_h}{0}{\eassignstack{x_h}{1}}}{\hole}}
{\ewidenderef{\rho}{x}{\tptr{\tntarrayb{(b_l,b_h)}{\tau}}{c}}{\dot C}}

\inferrule[]
{e \notin dom(\rho)}
{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{m}}{\hole}}

\inferrule[]
{x_l,x_h = \rho(e) \\
  x_a = \fresh \\
  \dot C = \elet{x_a}{\eif{\dot a \leq x_h}{0}{\eassignstack{x_h}{\dot a}}}{\hole}}
{\ewidenstrlen{\rho}{e}{\dot a}{\tptr{\tntarrayb{\bvar}{\tau}}{\cmode}}{\dot C}}

}
 \caption{Metafunctions for Widening}
\label{fig:compilationhelpers2}
\end{figure*}

\begin{figure*}[h]
{\small
  \begin{mathpar}
    \inferrule[]
    {x = \fresh \\
      \dot C = \elet{x}{\eif{\dot a}{0}{\enull}}{\hole}}
    {\echecknull{\dot a}{c}{\dot C}}

    \inferrule[]
    {}
    {\echecknull{\dot a}{u}{\hole}} \\

    \inferrule[]
    {}
    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}

    \inferrule[]
    { x_l,x_h = \rho(e) \\
      x_{cl},x_{ch} = \fresh\\
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
    }
    {\echeckbounds{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}

    \inferrule[]
    { e \notin dom(\rho) \\
      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
      \dot C_l = \elet{x_l}{b_l}{\hole}\\
      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
    }
    {\echeckbounds{\rho}{e}{\tntarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}

    \inferrule[]
    { e \notin dom(\rho) \\
      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
      \dot C_l = \elet{x_l}{b_l}{\hole}\\
      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
    }
    {\echeckbounds{\rho}{e}{\tarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}

    \inferrule[]
    {}
    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{u}}{\dot a}{\hole}}

    \inferrule[]
    { x_l,x_h = \rho(e) \\
      x_{cl},x_{ch} = \fresh\\
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{\dot a \leq x_h}{0}{\ebounds}}{\hole}
    }
    {\echeckboundsw{\rho}{e}{\tptr{\tallarrayb{\bvar}{\tau}}{c}}{\dot a}{\dot C_{cl;ch}}}

    \inferrule[]
    { e \notin dom(\rho) \\
      x_{l}, x_{h}, x_{cl},x_{ch} = \fresh\\
      \dot C_l = \elet{x_l}{b_l}{\hole}\\
      \dot C_h = \elet{x_h}{b_h}{\hole}  \\    
      \dot C_{cl} = \elet{x_{cl}}{\eif{x_l \leq \dot a}{0}{\ebounds}}{\hole}\\
      \dot C_{ch} = \elet{x_{ch}}{\eif{x_h \leq \dot a}{\ebounds}{0}}{\hole}
    }
    {\echeckboundsw{\rho}{e}{\tallarrayptr{b_l}{b_h}{\tau}{c}}{\dot a}{\dot C_{l;h;cl;ch}}}

\inferrule[]
{e \notin dom(\rho) \\
  x_l,x_l',x_h,x_h' = \fresh \\
  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
  \dot C_2 = \elet{x_l'}{b_l'}{\elet{x_h'}{b_h'}{\hole}}\\
  \dot C_3 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
  \dot C_4 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
}
{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3;4}}}

\inferrule[]
{x_l',x_h' =\rho(e) \\
  x_l,x_h = \fresh \\
  \dot C_1 = \elet{x_l}{b_l}{\elet{x_h}{b_h}{\hole}}\\
  \dot C_2 = \eif{x_l' \leq x_l}{\hole}{\ebounds} \\
  \dot C_3 = \eif{x_h \leq x_h'}{\hole}{\ebounds}
}
{\echeckboundsdyn{\rho}{e}{\tptr{\tallarrayb{(b_l,b_h)}{\tau}}{m}}{\tptr{\tallarrayb{(b_l',b_h')}{\tau}}{m}}{\dot C_{1;2;3}}}
  \end{mathpar}
}
\caption{Metafunctions for Dynamic Checks}
\label{fig:compilationhelpers}
\end{figure*}

% the closure $\dot
% C_{1;2;n;3;b;4}$ encodes the idea that $e_1$ and $e_2$ are evaluated to values
% before a null check (included in $\dot C_n$) is performed on
% $\dot a_1$ , the result of evaluating $e_1$. $\dot C_3$ adds $\dot
% a_1$ and $\dot a_2$ before a bounds check (included in $\dot C_b$) is performed on $\dot a_1$
% with $\dot a_2$ as the index. Finally, we use $\dot C_4$ 
% to store the result of the dereference in $x_4$ and return the pair
% $\dot C_4,x_4$ as the result of the compilation. 
% Further checks and
% widening can be performed using $x_4$ and we don't have to worry about
% the effects being triggered multiple times because the evaluation only
% happens once in $\dot C_4$.

% insert runtime checks and bounds widening
% code in a straightforward way. Take the \textbf{C-Deref} rule as an example,
% if we were to output $\dot e$ directly, we would have to rebind the
% expression to a variable before we can perform the null check and
% bounds check since $\dot e$ is not always a pure expression. Adding
% $\elettext$ bindings in such an ad-hoc fashion for avoiding side
% effects can clutter the compilation rules. Our approach of
% outputting a context and an atom threads the ANF transformation into
% each individual rule. Fig.~\ref{fig:compilationhelpers} shows helper
% functions that take atom expressions and return a context that
% contains the checks. The advantage of using explicit contexts becomes
% evident when we start to compose them. For the simulation property to
% hold, we want to ensure that the effects happen in the exact same
% order.  In
% the C-Deref rule, we first evaluate the expression that corresponds to
% $e$ through $\dot C_1$, and then perform a bounds check followed by a
% null check through $\dot C_b$ and $\dot C_n$. Finally, we dereference
% the expression through $\dot C_2$, potentially widen the upper bound
% by checking the dereferenced
% result.

% One big difference is that
% our compilation strategy is not parameterized by the heap. Given a
% literal annotated with a type, the compilation judgment simply
% discards the type without validating whether the literal's type is
% valid. This kind of relaxation not only makes the compilation rules
% simpler (we no longer need to keep track of $\defscope$ and the mode),
% but also allows us to formulate a stronger simulation theorem that
% applies even after the program state is corrupted. More concretely,
% the compilation judgment satisfies preservation regardless of whether
% an unchecked step is being taken. On the other hand, if we were to
% embed the compilation into the full-fledged typechecking rules, we
% would not always be able to compile the expression after a step has
% been taken.

% When our source
% program is stuck because of the execution of unchecked code, we can
% conclude that the compiled program will also get stuck with the
% same erased heap state. It is possible to embed our compilation
% strategy directly into the typing judgment, but our simulation would
% be limited to states that are not yet corrupted to be untypable.

% Since the erased language does not have type annotations, we need
% to explicitly represent bounds of null-terminated arrays as stack
% variables to support bounds widening at runtime. Every time a checked
% null-terminated array pointer is bound to a variable, we generate two
% fresh new variables that are assigned the lower and upper bounds of
% the array. The upper bound can be mutated into
% a larger bound during execution. We use $\rho$ to keep track of the
% bounds variables that belong to null-terminated array variables.

>>>>

\section{Appendix}\label{app:main}

@@

\subsection{The Compilation Rules}\label{appx:comp1}

% \review{
% The first reviewer said:
% \begin{itemize}
% \item I had a really hard time understanding precisely the "formalized" compilation
%   from CoreChkC to CoreC. Specifically: is CoreC intended to model LLVM-IR, or
%   is it a subset of C? Is CheckedC compiling to C or is it a new frontend like
%   `clang`? See remark about undefined behavior (UB) below which got me even more
%   confused. \mwh{CoreC is meant to model CoreChkC but with
%   annotations, and checks removed; it is not meant to model C or LLVM-IR.}
% \item Is the compilation scheme from CoreChkC to CoreC is faithful to the actual
%   compilation scheme of the checkedc-clang compiler to ...? I feel like the
%   paper is missing an actual description of what happens in the compiler to
%   allow us to connect the dots and understand how the formalization illuminates
%   the implementation. \mwh{We did not model compilation on the real
%   implementation; our purpose was to show that annotations in CoreChkC
%   do not necessitate fat pointers in an implementation; that said,
%   our formalization does show how a real implementation can be
%   carried out}
% \item The paper doesn't seem upfront about what is *shown* (theorem in Coq) and what
%   is *tested* (via PLT-Redex), and thus remains a
%   conjecture. \mwh{Updated the intro and the individual sections}
% \item  Missing discussion of why Coq vs. PLT-Redex, effort involved, any plans to
%   formally prove compilation from CoreChkC to CoreC, any hopes of integrating
%   that in the official implementation, etc. \mwh{Added note at the
%   start of section 3.}
% \end{itemize} }

% \review{IV: ghost variables in other contexts (e.g. Why3, Dafny) are used for things
%   that do not exist at run-time, but this doesn't seem to be the case here.}
% \yiyun{Agreed: We changed the name to ``shadow variables'' to avoid confusion}
% \review{Do your bug report and github links break anonymity?}

% \review{From reviewer C:
% \begin{itemize}
% \item I'd like to have seen a bit more motivation for using PLT redex:
%   what aspects made the use of this tool preferable to formulating the
%   compilation in Coq and using Quickchick to do random testing of the
%   simulation result.
%   \mwh{Added some text to the end of III.A and IV.C}
% \yiyun{Some reasons I can think of:\begin{itemize}
%     \item Redex is highly optimized for specifying judgments that are algorithmic. By writing down a typing relation, we can immediately obtain a typechecker
%     that is executable. Same applies for the small-step evaluation relation. Translating the relations into functions in Coq is definitely doable but time-consuming, especially since compilation is embedded as part of the typing rules. It is also hard to see whether the function we define really corresponds to the relation unless we formally prove it. This issue is particularly relevant at the early stage of the development when the compilation rules were buggy and the simulation property was violated often as we added new generator cases.
%     In Redex, we don't have any formal guarantee either, but at least we can more easily see the correspondence because Redex is able to convert the relation into an executable version so we can specify the relation literally. This feature of Redex helped us speed up our development significantly when our compilation rules were constantly changing.
% \end{itemize} }
% \item there's a funny change of line spacing in column 2 of page 6, about 2/3 down.
% \end{itemize}
% }

The main subtlety of compiling \checkedc to Clang/LLVM is to capture the annotations on pointer literals
that track array bound information, which is used in premises
of rules like \textsc{S-DefArray} and
  \textsc{S-AssignArr} to prevent spatial safety violations.
The \checkedc compiler \cite{li22checkedc} inserted additional pointer checks 
for verifying pointers are not null and the bounds are within their limits. \lc{what does it mean by bounds within limits?}
The latter is done by introducing additional shadow variables for storing (NT-)array pointer bound information.

\begin{figure}[t!]
{\small
\hspace*{-0.5em}
\begin{tabular}{|c|c|c|c|}
\hline
& \cmode & \tmode & \umode \\
\hline
& \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} & \textsc{CBox} / \textsc{Core} \\
\hline
\cmode & $\estar{x}$ / $\getstar{\cmode}{x}$ 
 & $\texttt{sand\_get}(x)$ / $\getstar{\umode}{x}$ &  $\times$ \\
\hline
\umode & $\times$
 & $\estar{x}$ / $\getstar{\umode}{x}$ &  $\estar{x}$ / $\getstar{\umode}{x}$ \\
\hline
\end{tabular}

}
\caption{Compiled Targets for Dereference}
\label{fig:flagtable}
\end{figure}

In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions. \lc{context mode doesn't determine where pointer points to but check if it is valid to point to that region?}
Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different. \lc{should we explain why we use different syntax instead of compiling the corresponding heap to llvm}
\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang). The columns represent different pointer modes and the rows represent context modes.
For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation. In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$. If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.

This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
with no loss of expressiveness
as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
For the compiler formalism, 
we present a compilation algorithm that converts from
\lang to \elang, an untyped language without metadata
annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$. \lc{strlen?}
The algorithm sheds
  light on how compilation can be implemented in the real Checked C
  compiler, while eschewing many vital details (\elang has many 
  differences with LLVM IR).

%This section shows how \systemname deals with 
%annotations can be safely erased: using static information a compiler
%can insert code to manage and check bounds metadata, with no loss of
%expressiveness. We present a compilation algorithm that converts from
%\lang to \elang, an untyped language without metadata
%annotations. The syntax and semantics \elang
  %closely mirrors that of \lang; it differs only in that literals lack
  %type annotations and its operational rules perform no
  %bounds and null checks, which are instead inserted during
  %compilation. Our compilation algorithm is evidence that \lang's
  %semantics, despite its apparent use of fat pointers, faithfully
  %represents Checked C's intended behavior. The algorithm also sheds
  %light on how compilation can be implemented in the real Checked C
  %compiler, while eschewing many important details (\elang has many 
  %differences with LLVM IR).

Compilation is defined by extending \lang's
typing judgment as follows:
\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
There is now a \elang output $\dot e$ and an input $\rho$, which maps
each (NT-)array pointer variable to its mode and
each variable \code{p} to a pair of \emph{shadow
  variables} that keep \code{p}'s up-to-date upper and lower bounds. 
These may differ from the bounds in \code{p}'s type due to bounds
widening.\footnote{Since lower bounds are never widened, the
  lower-bound shadow variable is unnecessary; we include it for uniformity.} 
\lc{These two sentences are confusing. }

% When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
% complete judgment, implicitly assuming that $e$ is a well-typed and closed
% term.

We formalize rules for this judgment in PLT Redex~\cite{pltredex},
following and extending our Coq development for \lang. To give
confidence that compilation is correct, we use Redex's property-based
random testing support to show that compiled-to $\dot e $ simulates
$e$, for all $e$.

% We developed a \checkedc compiler to compile a \checkedc program to a C program.
% \mwh{We formalized compilation from CoreChkC to a version of CoreChkC
%   but with the metadata removed, right? This is not a Checked C
%   compiler. You go on to see stuff about CompCert, CLight,
%   etc. This is confusing. We should be talking about how this relates
%   to what was just presented. Pick definitive names for things. }
% Given a \checkedc program $e$, we build a compilation process ($\gg$), such that $e \gg \dot e$, where $\dot e$ is the corresponding C program for $e$ in A-normal form (ANF). 
% The compilation process ($\gg$) relies on the type checking step $\Gamma;\Theta\vdash_m e:\tau$. 
% Especially, it relies on $\Gamma$ to provide the type information for variables in $e$. 
% We utilize CompCert/CLight syntax and semantics \cite{Leroy:2009:FVC:1666192.1666216,Blazy2009} as our translation target language.
%  Besides, we defined two data structures in the CompCert format for representing $\enull$ and $\ebounds$ states.
% We write $\xrightarrow{c}$ for the semantics of CLight. \mwh{Of our
%   target language, which I presume does not have all the features that
%   CLight has. Should mention up front that we did all of this in PLT Redex.}

Due to space constraints, we explain the rules for compilation by
examples, using a C-like syntax; the complete rules are given in
\iftr
Appendix~\ref{appx:comp1}.
\else
the supplemental report~\cite{checkedc-tech-report}.
\fi
Each rule performs up to three tasks: (a) conversion of $e$ to
A-normal form; (b) insertion of dynamic checks and bound widening expressions; 
and (c) generate right pointer accessing expressions based on modes.
%
A-normal form conversion is straightforward: compound expressions are managed by storing results of subexpressions into temporary variables,
as in the following example.

{\vspace*{-0.5em}
{\small
\begin{center}
$
\begin{array}{l}
$\code{let y=(x+1)+(6+1)}$
\;
\begin{frame}

\tikz\draw[-Latex,line width=2pt,color=orange] (0,0) -- (1,0);

\end{frame}
\;
\begin{array}{l}
$\code{let a=x+1;}$\\
$\code{let b=6+1;}$\\
$\code{let y=a+b}$\\
\end{array}
\end{array}
$
\end{center}
}
}

This simplifies the management of effects from subexpressions. The
next two steps of compilation are more interesting.
We state them based on different \lang operations.

% \review{Fig 9: if this is actual C code, then your null-check at line 6 will be
%   eliminated by the compiler. At line 3, you performed a pointer addition, which
%   is only defined when `p` is non-null. So, either `p` is non-null, and the
%   NULL-check can be eliminated; or, `p` is NULL, but line 3 was undefined
%   behavior, meaning the compiler is allowed to do anything, notably eliminate
%   the NULL-check. This is where I am super confused, and either:
%   - CoreC is not really the C language, and has different semantics...? but is
%     this well-defined in the context of LLVM?
%   - there is a problem that was not caught by the PLT-Redex-based testing.
% \yiyun{We have clarified at the start of IV that CoreC is an untyped
%   variant of CoreChkC, and does not aim to represent C per se, or LLVM
%   IR. We aimed to avoid confusion by rewriting the examples in a way that is more closely
%     related to the syntax presented in Fig 3. Pointer arithmetic between 0 and a non-zero
%     index is always valid because CoreC there is technically only
%     integer arithmetic.}}

\myparagraph{Pointer Accesses and Modes}
%
In every declaration \lc{(or the beginning of a function body) why?} of a pointer,
if the poniter is an (NT-)array,
we first allocate two \emph{shadow variables}
to track the lower and upper bounds which are potentially changed for pointer arithmetic and NT-array bound widening.
Each $\cmode$-mode NT-array pointer variable is associated with its type information in a store.
Additionally, we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
there is an inserted verification step similar to \Cref{fig:const-type},
which checks if a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
 and assignment (\code{*(mode(q) }$\wedge$\code{ m,q)=1}) operations 
are computed based on the meet 
operation ($\wedge$) of the pointer mode (e.g. \code{mode(p)}) and the current context mode (\code{m}).

\myparagraph{Checked and Unchecked Blocks}
%
In the \systemname implementation,
$\euncheckedtext$ and $\echeckedtext$ blocks 
are compiled as context switching functions provided by the sandbox mechanism.
$\eunchecked{\overline{x}}{e}$ is compiled to 
$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
to execute expression $e$ with the arguments $\overline{x}$.
$\echecked{\overline{x}}{e}$ is compiled to 
$\texttt{callback}(\overline{x},e)$, where we perform 
a \texttt{callback} to a checked block code $e$ inside a sandbox.
In \systemname, we adopt an aggressive execution scheme that
directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
In the formalism, we rely on the type system to 
guarantee the context switching without creating the extra function calls for simplicity.

%Fig.~\ref{fig:compilationexample} shows how an invocation of
%\code{strlen} on a null-terminated string is compiled into C
%code. Each dereference of a checked pointer requires a null check
%(See \textsc{S-DefNull} in Fig.~\ref{fig:semantics}), which the
%compiler makes explicit: Line~$3$ of the generated code has the null
%check on pointer \code{p} due to the \code{strlen},
%  and a similar check happens
%  at line~$8$ due to the pointer arithmetic on \code{p}.
%Dereferences also require bounds checks: line~$2$ checks \code{p} is
%in bounds before computing \code{strlen(p)}, while line~$10$ does
%likewise before computing \code{*(p+1)}.

\myparagraph{Function Pointers and Calls}
%
Function pointers are managed similarly to normal pointers,
but we insert checks to check if the pointer address is not null in 
the function store instead of heap, and whether or not the type is correctly represented, 
for both $\cmode$ and $\tmode$ mode pointers 
\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.

For compiling dependent function calls,
\Cref{fig:compilationexample1} provides a hint.
Notice that the bounds for the array pointer \code{p} are not passed as
arguments. Instead, they are initialized according to \code{p}'s
type---see line~4 of the original \lang program at the top of the figure.
Line~$3$ of the generated code
sets the lower bound  to \code{0} and the
upper bound to \code{n}.

\subsection{Metatheory}
\label{sec:meta}

% \review{ 
% My interpretation of section IV.C is that the authors have a pen-and-paper
%   proof of Theorem 4, and that random testing using the Redex models has been
%   used to gain confidence in such a proof. Is this correct?}
% \liyi{Yes, this is correct.}
% \mwh{No, not correct: We have no pen-and-paper proof, just the
%   model. Clarify here}
% showing that
% extensive random testing fails to falsify the bisimilarity
% property. \mwh{Better way to state the previous?}

We formalize both the compilation procedure and the simulation
theorem in the PLT Redex model we developed for \lang (see Sec.~\ref{sec:syntax}),
and then attempt to falsify it via Redex's support for random
testing. Redex allows us
  to specify compilation as logical rules (an extension
  of typing), but then execute it algorithmically to
  automatically test whether simulation holds. This process revealed
  several bugs in compilation and the theorem statement.
%
  % us gain confidence that our original pen and paper proof of
  % simulation remains true with the addition of variable bounds. }
We ultimately plan to prove simulation in the Coq model.

%Turning to the simulation theorem: We first introduce notation
%used to specify the theorem.
We use the notation $\gg$ to
indicate the \emph{erasure} of stack and heap---the rhs is the same as
the lhs but with type annotations removed:
\begin{equation*}
  \begin{split}
    \heap  \gg & \dot \heap \\
    \varphi \gg & \dot \varphi
  \end{split}
\end{equation*}
In addition, when $\Gamma;\emptyset\vdash
\varphi$ and $\varphi$ is well-formed, we write $(\varphi,\heap,e) \gg_m (\dot \varphi, \dot \heap,
\dot e)$ to denote $\varphi \gg \dot \varphi$, $\heap \gg \dot \heap$
and $\Gamma;\Theta;\emptyset \vdash_m e \gg \dot e : \tau$ for some $\tau$ respectively. $\Gamma$ is omitted from the notation since the well-formedness of $\varphi$ and its consistency with respect to $\Gamma$ imply that $e$ must be closed under $\varphi$, allowing us to recover $\Gamma$ from $\varphi$.
Finally, we use $\xrightarrow{\cdot}^*$ to denote the transitive closure of the
reduction relation of $\elang$. Unlike the $\lang$, the semantics of
$\elang$ does not distinguish checked and unchecked regions.

Fig.~\ref{fig:checkedc-simulation-ref} gives an overview of 
the simulation theorem.\footnote{We ellide the  possibility of $\dot e_1$ evaluating to $\ebounds$ or $\enull$ in the diagram for readability.} The simulation theorem is specified in a way
that is similar to the one by~\citet{merigoux2021catala}.

An ordinary simulation property would
replace the middle and bottom parts of the figure with the
following: \[(\dot \varphi_0, \dot \heap_0, \dot e_0) 
  \xrightarrow{\cdot}^* (\dot \varphi_1, \dot \heap_1, \dot e_1)\]
Instead, we relate two erased configurations using the relation $\sim$,
which only requires that the two configurations will eventually reduce
to the same state.
